A Classical Density-Functional Theory for Describing Water Interfaces

### 2012-12-13 papers/water-SAFT/paper.tex

\author{Jessica Hughes}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}
\author{Eric Krebs}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms-truncated}
\end{center}
\caption{The pressure versus density for various temperatures, including
experimental pressure data from NIST\cite{nistwater}. The solid colored lines
indicate the computationally calculated pressure and the dotted
colored lines are NIST data points. The solid and dotted black lines
represent the theoretical and experimental coexistence curves.}
\label{fig:pressure-with-isotherms}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/surface-tension}
\end{center}
\caption{Comparison of Surface tension versus temperature for theoretical and
  experimental data. The experimental data is taken from NIST.\cite{nistwater}
  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
  tension will match the experimental surface tension near room temperature.}
\label{fig:surface-tension}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
\end{center}
\caption{ Excess chemical potential per area versus radius for a single
hydrophobic rod immersed in water. This should have an asymptote equal
to the surface tension at room temperature, and it agrees well with
the surface tension in Figure~\ref{fig:surface-tension}. The inset for
rods with a very small radius shows the linear relationship expected based on
Equation~\ref{contactvaluethm}.}
\label{fig:energy-vs-diameter}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-single-rod}
\end{center}
\caption{ Density profiles for single rods of different radii. The dotted line
represents the saturated liquid density and the points represent the
expected contact density derived from the contact value theorem and
calculated free energy data.}
\label{fig:density-single-rod}
\end{figure}

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=\columnwidth]{figs/rods-diagram}
%\end{center}
%\caption{ Cross section schematic for two hydrophobic rods immersed in water.
%The radius and separation are shown.}
%\label{fig:rods}
%\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
\end{center}
\caption{ Density profiles illustrating the transition from vapor 
to liquid water between the rods. The radius is 0.6~nm, the top figure is
at a separation of 0.6~nm and the
bottom is 0.7~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
the energy for these and other separations.}
\label{fig:density-rods}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
\end{center}
\caption{ Free energy of interaction (also known as the potential of mean force)
versus separation for two hydrophobic rods ranging in radius from
0.2~nm to 1.2~nm.
All were arbitrarily offset to zero at large separations for ease of comparison. The
transition corresponds to the phase change from
vapor to liquid between the rods as pictured in the density profiles in 
Figure~\ref{fig:density-rods}. }
\label{fig:rods-energy-vs-distance}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/four-rods-energy-vs-distance}
\end{center}
\caption{ Free energy of interaction
versus separation for four hydrophobic rods ranging in radius from
0.2~nm to 1.2~nm.
All were arbitrarily offset to zero at large separations. The
transition corresponds to the phase change from
vapor to liquid between the rods as pictured in the density profiles in 
Figure~\ref{fig:density-four-rods}. }
\label{fig:four-rods-energy-vs-distance}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-four-rods}
\end{center}
\caption{ Density profiles illustrating the transition from vapor 
to liquid water between four rods. The radius is 0.6~nm, the top figure is
at a separation of 1.53~nm and the
bottom is 1.56~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
the energy for these and other separations.}
\label{fig:density-four-rods}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/sphere-energy-vs-diameter}
\end{center}
\caption{ Excess chemical potential per area versus radius for
  a single hydrophobic sphere
  immersed in water. This should have an asymptote equal to the
  surface tension at room temperature, and it agrees well with the
  surface tension in Figure~\ref{fig:surface-tension}. Results from a
  simulation of SPC/E water~\cite{huang2001shs} are shown as circles.
  The horizontal lines show the experimental and SPC/E macroscopic surface
  tension for water at standard atmospheric temperature and
  pressure. }
\label{fig:sphere-energy-vs-diameter}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-sphere}
\end{center}
\caption{ Density profiles around hard-sphere solutes of different radii. Predictions
  from our classical density-functional theory are in solid red, while
  the dotted line shows the result of a molecular dynamics simulation
  of SPC/E water~\cite{huang2001shs}.  }
\label{fig:density-sphere}
\end{figure}

>>>>

\author{Jessica Hughes}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}
\author{Eric J. Krebs}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331, USA}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/pressure-with-isotherms-truncated}
\end{center}
\caption{The pressure versus density for various temperatures, including
experimental pressure data from NIST\cite{nistwater}. The solid colored lines
indicate the computationally calculated pressure and the dotted
colored lines are NIST data points. The solid and dotted black lines
represent the theoretical and experimental coexistence curves.}
\label{fig:pressure-with-isotherms}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/surface-tension}
\end{center}
\caption{Comparison of Surface tension versus temperature for theoretical and
  experimental data. The experimental data is taken from NIST.\cite{nistwater}
  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
  tension will match the experimental surface tension near room temperature.}
\label{fig:surface-tension}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/energy-vs-diameter}
\end{center}
\caption{ Excess chemical potential per area versus radius for a single
hydrophobic rod immersed in water. This should have an asymptote equal
to the surface tension at room temperature, and it agrees well with
the surface tension in Figure~\ref{fig:surface-tension}. The inset for
rods with a very small radius shows the linear relationship expected based on
Equation~\ref{contactvaluethm}.}
\label{fig:energy-vs-diameter}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-single-rod}
\end{center}
\caption{ Density profiles for single rods of different radii. The dotted line
represents the saturated liquid density and the points represent the
expected contact density derived from the contact value theorem and
calculated free energy data.}
\label{fig:density-single-rod}
\end{figure}

@@

%\begin{figure}
%\begin{center}
%\includegraphics[width=3.5in]{figs/rods-diagram}
%\end{center}
%\caption{ Cross section schematic for two hydrophobic rods immersed in water.
%The radius and separation are shown.}
%\label{fig:rods}
%\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-rods-in-water}
\end{center}
\caption{ Density profiles illustrating the transition from vapor 
to liquid water between the rods. The radius is 0.6~nm, the top figure is
at a separation of 0.6~nm and the
bottom is 0.7~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
the energy for these and other separations.}
\label{fig:density-rods}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/rods-energy-vs-distance}
\end{center}
\caption{ Free energy of interaction (also known as the potential of mean force)
versus separation for two hydrophobic rods ranging in radius from
0.2~nm to 1.2~nm.
All were arbitrarily offset to zero at large separations for ease of comparison. The
transition corresponds to the phase change from
vapor to liquid between the rods as pictured in the density profiles in 
Figure~\ref{fig:density-rods}. }
\label{fig:rods-energy-vs-distance}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/four-rods-energy-vs-distance}
\end{center}
\caption{ Free energy of interaction
versus separation for four hydrophobic rods ranging in radius from
0.2~nm to 1.2~nm.
All were arbitrarily offset to zero at large separations. The
transition corresponds to the phase change from
vapor to liquid between the rods as pictured in the density profiles in 
Figure~\ref{fig:density-four-rods}. }
\label{fig:four-rods-energy-vs-distance}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-four-rods}
\end{center}
\caption{ Density profiles illustrating the transition from vapor 
to liquid water between four rods. The radius is 0.6~nm, the top figure is
at a separation of 1.53~nm and the
bottom is 1.56~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
the energy for these and other separations.}
\label{fig:density-four-rods}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/sphere-energy-vs-diameter}
\end{center}
\caption{ Excess chemical potential per area versus radius for
  a single hydrophobic sphere
  immersed in water. This should have an asymptote equal to the
  surface tension at room temperature, and it agrees well with the
  surface tension in Figure~\ref{fig:surface-tension}. Results from a
  simulation of SPC/E water~\cite{huang2001shs} are shown as circles.
  The horizontal lines show the experimental and SPC/E macroscopic surface
  tension for water at standard atmospheric temperature and
  pressure. }
\label{fig:sphere-energy-vs-diameter}
\end{figure}

@@

\begin{figure}
\begin{center}
\includegraphics[width=3.5in]{figs/density-sphere}
\end{center}
\caption{ Density profiles around hard-sphere solutes of different radii. Predictions
  from our classical density-functional theory are in solid red, while
  the dotted line shows the result of a molecular dynamics simulation
  of SPC/E water~\cite{huang2001shs}.  }
\label{fig:density-sphere}
\end{figure}
